---
title: "HW2"
author: "Mikayla Norton"
date: "2023-01-21"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(sqldf)
```

# Problem 1
The purpose of this problem is to create a modeling dataset for customer churn in the next year, based on the customer history dataset. In ICA 3, you performed the row augmentation to get missing rows added to the data. To get to the modeling dataset:

## Part A
Create the target field based on the strategy given in the slides.
```{r p1a}
customer_history <- read.csv("orders/Customer_History.csv")
years <- data.frame(year = seq(2005,2022))

customer_history2 <- sqldf("SELECT *
               FROM customer_history
               INNER JOIN years
               WHERE year <= Last_Year 
               AND year >= First_Year")
head(customer_history2)
```

## Part B
Create the following additional input variables that would be used to make predictions:
i. Customer age
```{r p1bi}
customer_history2 <- sqldf("SELECT *, year-Year_born as Age
               FROM customer_history2")

head(customer_history2)
```
ii. Number of years as a customer
```{r p1bii}
customer_history2 <- sqldf("SELECT *, Last_Year-First_Year as Years_As_Cust
               FROM customer_history2")
head(customer_history2)
```

iii. Using the complaints dataset, create a count of complaints made each year and join it to the dataset (be careful with what year you join!)
```{r p1biii}
complaints <- read.csv("orders/Complaint_History.csv")

complaintsSummary <- sqldf("SELECT CustomerID, ComplaintYear,COUNT(*) as ComplaintCount
                           FROM complaints
                           GROUP BY ComplaintYear, CustomerID
                           ORDER BY ComplaintYear")

Cust_Complaints <- sqldf("SELECT customer_history2.*,ComplaintCount
                         FROM customer_history2
                         LEFT JOIN complaintsSummary
                         ON customer_history2.year = complaintsSummary.ComplaintYear 
                         AND customer_history2.Customer=complaintsSummary.CustomerID")
Cust_Complaints["ComplaintCount"][is.na(Cust_Complaints["ComplaintCount"])] <- 0

head(Cust_Complaints)
```


# Problem 2
The purpose of this problem is to create a modeling dataset for predicting which orders will order which products, based on the order history dataset. In ICA 3, you performed the row augmentation. Now:

## Part A
Create the target field based on the strategy given in the slides.
```{r p2a}
order_history <- read.csv("orders/Order_History.csv")
order_history2 <- sqldf("SELECT CustomerID, Product_ID, Year, Month, sum(Quantity) as Quantity
                        FROM order_history
                        GROUP BY customerID, Product_ID, Year, Month
                        ORDER BY CustomerID, Product_ID, Year, Month")

order_wide <- pivot_wider(order_history2, names_from = Year, values_from = Quantity)
order_long <- pivot_longer(order_wide, cols = starts_with('20'), names_to = 'Year', values_to = 'Quantity',)
order_long <- sqldf("SELECT *
                    FROM order_long
                    ORDER BY  CustomerID, Product_ID, Year, Month")

order_long["Quantity"][is.na(order_long["Quantity"])] <- 0
order_long["Quantity"][order_long["Quantity"] != 0] <- 1

order_long <- order_long %>% 
       rename("OrderBinary" = "Quantity")

head(order_long)
```
```{r p2-additional setup}
year <- data.frame("year" = min(order_long$Year):max(order_long$Year))
month <- data.frame("month" = min(order_long$Month):max(order_long$Month))
cust <- data.frame("customer" = min(order_long$CustomerID):max(order_long$CustomerID))
products <- data.frame("product" = min(order_long$Product_ID):max(order_long$Product_ID))

cartesian_orders <- sqldf("SELECT *
                   FROM year
                   INNER JOIN month
                   INNER JOIN cust
                   INNER JOIN products")

# order_summary <- sqldf("SELECT *, 
#                     CASE WHEN Month = 1 THEN 12 ELSE Month-1 END AS PastMonth
#                     FROM order_long
#                     GROUP BY CustomerID, Year, Month")
# 
# order_summary <- sqldf("SELECT *,
#                     CASE WHEN Month = 1 THEN Year-1 ELSE Year END AS PastYear
#                     FROM order_summary
#                     GROUP BY CustomerID, Year, Month")
# head(order_summary)
```


## Part B
Create the following additional input variables that would be used to make predictions. All of these x variables should be from the standpoint of the prior month.

i. Number of months since the last order
```{r p2bi}

Order_summaries <- sqldf("SELECT customer, product, t1.year, t1.month, SUM(OrderBinary) as TotalOrders
                         FROM cartesian_orders t1
                         LEFT JOIN order_long t2
                         ON t1.customer=t2.CustomerID 
                         AND t1.product=t2.Product_ID 
                         AND t1.year=t2.Year 
                         AND t1.month=t2.Month
                         GROUP BY customer, product, t1.year, t1.month")

Order_summaries["TotalOrders"][is.na(Order_summaries["TotalOrders"])] <- 0

#####


Months_Since <- sqldf("SELECT t1.month, t1.year, t1.customer,t1.product, t1.TotalOrders, (t1.year*12+t1.month)-(t2.year*12+t2.month) as MonthsSinceLast
                         FROM Order_summaries t1
                         INNER JOIN Order_summaries t2
                         ON t1.customer=t2.customer
                         AND t1.product=t2.product
                         WHERE t1.TotalOrders>0 AND t2.TotalOrders>0 AND MonthsSinceLast > 0 AND MonthsSinceLast <= 12")

Months_Since_Last <- sqldf("SELECT customer, product, month, year, min(MonthsSinceLast) as MonthsSinceLast
                         FROM Months_Since
                         GROUP BY customer, product, year, month
                         ORDER BY customer, product, year, month")
head(Months_Since_Last)
```

ii. Number of months out of the previous 12 months that an order was placed
```{r p2bii}
Previous12 <- sqldf("SELECT customer, product, month, year, COUNT(MonthsSinceLast) as OrdersOverLastYear
                    FROM Months_Since 
                    GROUP BY customer, product, year, month
                    ORDER BY customer, product, year, month")
head(Previous12)
```


iii. Average quantity ordered per order (over last 12 years)
```{r p2biii}

Avg_Quantity <- sqldf("SELECT t1.Month, t1.Year, t1.CustomerID, t1.Product_ID, avg(t2.Quantity) as AvgQuantity,  (t1.year*12+t1.month)-(t2.year*12+t2.month) as MonthsSinceLast
                         FROM order_history2 t1
                         INNER JOIN order_history2 t2
                         ON t1.CustomerID=t2.CustomerID
                         AND t1.Product_ID=t2.Product_ID
                         WHERE t1.Quantity>0 AND t2.Quantity>0 AND MonthsSinceLast >= 0 AND MonthsSinceLast <= 12
                         GROUP BY t1.CustomerID, t1.Product_ID, t1.Year, t1.Month
                         ORDER BY t1.CustomerID, t1.Product_ID, t1.Year, t1.Month")

Avg_Quantity <- sqldf("SELECT CustomerID, Product_ID, Month, Year, AvgQuantity
                      FROM Avg_Quantity
                      ORDER BY CustomerID, Product_ID, Year, Month")
head(Avg_Quantity)
```


# Problem 3
With the OJ dataset, calculate some additional X’s and join them back to the original dataset; specifically:
```{r p3}
oj <- ISLR2::OJ
head(oj)
```
## Part A
The average price for each brand by store
```{r p3a}
ojAvgPrice <- sqldf("SELECT *, avg(PriceCH) as AvgPriceCH, avg(PriceMM) as AvgPriceMM
             FROM oj
             GROUP BY StoreID")
head(ojAvgPrice)
```

## Part B
The previous week’s fraction of customers who bought Minute Maid, by store (note: you will lose the first week when you use this as an X)
```{r p3b}

ojCustCounts <- sqldf("SELECT *, COUNT(*) as TotalPurchases, 
                    COUNT(CASE WHEN Purchase = 'MM' THEN 1 ELSE NULL END) AS CountMM
                    FROM oj
                    GROUP BY WeekofPurchase")

ojCustCounts$TotalPurchases <- as.double(ojCustCounts$TotalPurchases)
ojCustCounts$CountMM <- as.double(ojCustCounts$CountMM)

ojFrac <- sqldf("SELECT t1.*, (t2.CountMM)/(t2.TotalPurchases) as CustFrac
                FROM ojCustCounts t1
                INNER JOIN ojCustCounts t2
                WHERE t1.WeekofPurchase = t2.WeekofPurchase+1")




head(ojFrac)
```

