---
title: "HW1"
author: "Mikayla Norton"
date: "2023-01-10"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Load data
library(sqldf)
library(tidyverse)
order_details <- read.csv("orders/order_details.csv")
orders <- read.csv("orders/orders.csv")
territories <- read.csv("orders/territories.csv")
regions <- read.csv("orders/regions.csv")
employee_territories <- read.csv("orders/employee_territories.csv")
employees <- read.csv("orders/employees.csv")
customers <- read.csv("orders/customers.csv")
shippers <- read.csv("orders/shippers.csv")
suppliers <- read.csv("orders/suppliers.csv")
products <- read.csv("orders/products.csv")
categories <- read.csv("orders/categories.csv")
```


# Problem 1
Perform a sort of orders by employeeID, then by shipVia, and then by freight, for those orders by shipped to France.
```{r}
ordersFrance <- sqldf("SELECT *
                      FROM orders
                      WHERE shipCountry is 'France'
                      ORDER BY employeeID, shipVia, freight")
head(ordersFrance)
```

# Problem 2
Which shipVia has the largest average cost?
```{r}
order_cost <- sqldf("SELECT order_details.*, unitPrice*quantity*(1-discount) AS cost 
                       FROM order_details")
```

```{r}
avgCost <- sqldf("SELECT shipVia, avg(cost) as AvgCost 
                 FROM orders 
                 INNER JOIN order_cost 
                 WHERE orders.orderID = order_cost.orderID
                 GROUP BY shipVia")
maxAvgCost <- sqldf("SELECT shipVia, max(AvgCost)
                    FROM avgCost")
maxAvgCost
```

# Problem 3
Which product category has the highest average UnitPrice? The Lowest?
```{r}
avgUnitPrice <- sqldf("SELECT CategoryID, avg(UnitPrice) as AvgUnitPrice
                      FROM products
                      GROUP BY CategoryID
                      ORDER BY avg(UnitPrice)")
```

```{r}
maxAvgPrice <- sqldf("SELECT CategoryID, max(AvgUnitPrice)
                     FROM avgUnitPrice")

minAvgPrice <- sqldf("SELECT CategoryID, min(AvgUnitPrice)
                     FROM avgUnitPrice")
maxAvgPrice
minAvgPrice
```

# Problem 4
Which products are supplied by a company in the United States?
```{r}
productUSA <- sqldf("SELECT DISTINCT ProductName 
                     FROM products 
                     INNER JOIN suppliers 
                     WHERE products.SupplierID = suppliers.SupplierID 
                     AND country='USA'")
head(productUSA)
```

# Problem 5
Which shipper is shipping the largest number of units of product? Answer in terms of units; you do not need to consider quantityPerUnit here.
```{r}
BigShipper <- sqldf("SELECT shipperID, sum(quantity) as UnitsShipped 
                     FROM orders 
                     INNER JOIN order_details 
                     INNER JOIN shippers
                     ON orders.orderID = order_details.orderID
                     AND shippers.shipperID = orders.shipVia
                     GROUP BY shipperID
                     ORDER BY UnitsShipped DESC
                     LIMIT 1")
BigShipper
```

# Problem 6
Which employee is tied to the most sales revenue? Give the name, not the code, along with the total revenue for the employee.
```{r}
maxEmployee <- sqldf("SELECT firstName, lastName, sum(unitPrice*quantity*(1-discount)) AS revenue 
                     FROM orders 
                     INNER JOIN order_details
                     INNER JOIN employees
                     ON orders.orderID = order_details.orderID
                     AND orders.employeeID = employees.employeeID
                     GROUP BY orders.employeeID
                     ORDER BY revenue DESC
                     LIMIT 1")
maxEmployee
```

# Problem 7
Find the total revenue for each product category.
```{r}
categorical_revenue <- sqldf("SELECT products.CategoryID, order_details.unitPrice*quantity*(1-discount) AS revenue 
                       FROM order_details
                       INNER JOIN products
                       INNER JOIN categories
                       ON order_details.productID = products.ProductID
                       AND products.CategoryID = categories.categoryID
                       GROUP BY products.CategoryID")

head(categorical_revenue)
```


# Problem 8
Consider the amount of revenue for each customer. If there were no discounts applied, which customer would see the largest increase in cost?
```{r}
CostInc <- sqldf("SELECT customerID, sum(unitPrice*quantity - unitPrice*quantity*(1-discount)) AS CostIncrease
                       FROM order_details
                       INNER JOIN orders
                       WHERE order_details.orderID = orders.orderID
                      GROUP BY customerID
                      ORDER BY CostIncrease DESC
                      LIMIT 1")
CostInc
```

# Problem 9
Which order(s) has the most number of items (and how many)? Give the orderID for this one.
```{r}
mostItems <- sqldf("SELECT orderID, sum(quantity) as totalItems
                   FROM order_details
                   GROUP BY orderID")
mostItems <- sqldf("SELECT orderID, max(totalItems)
                   FROM mostItems")
mostItems
```

# Problem 10
Create a new field called “InventoryOrderRatio” which is, for each product, the UnitsinStock (the inventory) for the product (across all customers) divided by the quantity ordered for that product. A high value represents sufficient product in stock, while a low number represents products that are in danger of running out. What 3 products are most in danger of running out?
```{r}
products2 <- sqldf("SELECT products.*, sum(quantity) as QuantityOrdered, UnitsinStock/sum(quantity) as InventoryOrderRatio
                  FROM products
                  INNER JOIN order_details
                  WHERE products.ProductID = order_details.productID
                  GROUP BY order_details.productID
                  ORDER BY InventoryOrderRatio
                  LIMIT 3")
products2
```

# Problem 11
A recommender engine looks at which pairs of products tend to be bought by the same customer, so that if a customer buys one, the recommender engine will recommend they buy the other. Find which product pairs are most likely to be bought by the same customer.
```{r}
library(sqldf)
rec <- sqldf("SELECT products.ProductID, ProductName, customers.customerID
             FROM products
             INNER JOIN customers
             INNER JOIN orders
             INNER JOIN order_details
             ON orders.orderID = order_details.orderID
             AND orders.customerID = customers.customerID
             AND order_details.ProductID = products.ProductID")


productsCustomers <- sqldf("SELECT v1.ProductName as Product1, v2.ProductName as Product2, count(DISTINCT v1.customerID) as totalCustomers
               FROM rec v1
               INNER JOIN rec v2
               ON v1.customerID = v2.customerID
               WHERE
                   v1.ProductID < v2.ProductID
               GROUP BY
                   v1.ProductName, v2.ProductName
               ORDER BY totalCustomers DESC")
head(productsCustomers)
```